### 2017-07-03 opsem.tex
RG Rules and explanation.

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in the \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a \R{TXN} block tagged with sets ($\stl$ and $\stg$) of
records whose meaning is explained below, and a runtime \R{FOREACH}
expression that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level syntax of the \C{FOREACH} command shown here is
a little different from the one used in previous sections; its
higher-order function has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning that will be useful for verification
(Sec.~\ref{sec:reasoning}).

>>>>

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in the \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a \textsc{TXN} block tagged with sets ($\stl$ and $\stg$) of
records whose meaning is explained below, and a runtime \textsc{FOREACH}
expression that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level syntax of the \C{FOREACH} command shown here is
a little different from the one used in previous sections; its
higher-order function has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning that will be useful for verification
(Sec.~\ref{sec:reasoning}).

### 2017-07-03 reasoning.tex
RG Rules and explanation.

>>>>

The \rulelabel{RG-Conseq} rule lets us safely strengthen the guarantee
$G$, or weaken the rely $R$ of a transaction. Importantly, it also
allows its isolation specification $\I$ to be strengthened. This means
that a transaction proven correct under a weaker isolation level is
also correct under a stronger level. Parametricity over the isolation
specification $\I$, combined with the ability to strengthen $\I$ as
needed, admits a flexible proof strategy to prove database programs
correct. For example, programmers can declare isolation requirements
of their choice through $\I$, and then prove programs correct assuming
the guarantees hold. The soundness of strengthening $\I$ ensures that
a program can be safely executed on any system that offers isolation
guarantees at least as strong as those assumed.

Two rules of the transaction-local RG judgment are shown in
Fig.~\ref{fig:rg-rules}. The rule \rulelabel{RG-Update} is
illustrative of the RG rules for SQL statements; they basically
reflect the structure of the corresponding reduction rule from
Fig.~\ref{fig:txnimp}, and require no further explanation. The rule
\rulelabel{RG-Foreach} defines the RG judgment for a \C{FOREACH} loop.
As is characteristic of loops, the reasoning is pivoted on a loop
invariant $\I$ (not same as the high-level invariant), that needs to
be stable w.r.t $\R$. $I$ must follow $P$, the pre-condition of
\C{FOREACH}, when no elements have been iterated, i.e, when
$y=\emptyset$. The body of the loop can assume the loop invariant, and
the fact that $z$ is an element from the set $x$ being iterated, to
prove its post-condition $Q_c$. Operational semantics ensures that $z$
is added to $y$ at the end of the iteration, hence $Q_c \conj z\in y$
is valid at the end of the iteration, from which $I$ must follow. When
\C{FOREACH} has finished execution, $y$, the set of iterated items, is
the entire set $x$. Thus $I \conj y=x$ must imply the post-condition
$Q$, which also needs to be stable. The rules for conditions,
sequencing etc., are more-or-less standard, hence elided.
